---
title: "TTF"
author: "Ziying Wang"
date: "12/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=F}
# library
library(readr)
library(tidyverse)
library(ggplot2)
# read data
data_full <- read_csv('Manu_data_200000.csv')
data_full<-data_full[,-c(1,17,19)]
```

# Just look at Denver Drivers as an example to simplify

```{r}
# 412 jobs
denver.driver <- filter(data_full, role %in% c("Truck Driver", "Driver", "CDL Driver"), region_state=="Denver-Aurora, CO MSA")
# Remove jobs only posted for 0 or 1 day
# 387 jobs, NA removed by default
denver.driver <- filter(denver.driver, time_to_fill > 1)
# Push the job id to the end
denver.driver$job_id_end <- denver.driver$job_id
denver.driver <- denver.driver[-c(1)]
denver.driver$region <- ifelse(denver.driver$city == "Denver", "Denver", ifelse(denver.driver$city == "Aurora", "Aurora", "Other"))
denver.driver$region <- as.factor(denver.driver$region)
denver.driver$region <- relevel(denver.driver$region, "Other")
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=salary, y=time_to_fill)) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = salary, y=time_to_fill, color=region), se=FALSE)
```

```{r}
# Try to predict time to fill based on salary, region = ("Other", "Denver", "Aurora")
den.driver.model <- lm(time_to_fill ~ salary + region, data=denver.driver)
summary(den.driver.model)
```

```{r}
# Remove extra high salaries (Many jobs have "accounting" in tags)
# 373 jobs, NA removed by default
denver.driver <- filter(denver.driver, salary < 100000)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=salary, y=time_to_fill)) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = salary, y=time_to_fill, color=region), se=FALSE)
```

```{r}
# Try to predict time to fill based on salary, region = ("Other", "Denver", "Aurora")
den.driver.model <- lm(time_to_fill ~ salary + region, data=denver.driver)
summary(den.driver.model)
# Inside / Outside Denver or Aurora or suberbs does not matter
```

# There was barely a relationship between "time_to_fill" and "salary" or "region = Denver, Aurora, Other" and when outliers removed, things are even worse. 

# Add indicator tags to determine if there are any relationships with the top driving tags:
# "Safety", "High School Diploma", "Training", "Transportation", "Customer Service", "Commercial", "Warehouse"

```{r}
denver.driver$safety <- as.numeric(str_detect(denver.driver$tags, "Safety"))
denver.driver$diploma <- as.numeric(str_detect(denver.driver$tags, "High School Diploma"))
denver.driver$training <- as.numeric(str_detect(denver.driver$tags, "Training"))
denver.driver$transportation <- as.numeric(str_detect(denver.driver$tags, "Transportation"))
denver.driver$cust.serv <- as.numeric(str_detect(denver.driver$tags, "Customer Service"))
denver.driver$commercial <- as.numeric(str_detect(denver.driver$tags, "Commercial"))
denver.driver$warehouse <- as.numeric(str_detect(denver.driver$tags, "Warehouse"))
```

```{r}
denver.driver$post_date<-as.Date(denver.driver$post_date,"%m/%d/%Y")
denver.driver$fill_date<-as.Date(denver.driver$fill_date,"%m/%d/%Y")
```

```{r include=F}
# Average time to fill driving jobs in Denver
# Month 1 = August 2017
# Month 12 = July 2018
library(lubridate)
```

```{r}
# Driver job postings by week in Denver
denver.driver %>%
  count(week = floor_date(post_date, "week")) %>%
  ggplot(aes(week, n)) + 
  geom_line() +
  labs(title = "Denver Driving Jobs (August 2017-July 2017), Job Posts Per Week")
```

```{r}
denver.driver$quarter1 <- ifelse(month(denver.driver$post_date) %in% c(1, 2, 3), 1, 0) 
denver.driver$quarter2 <- ifelse(month(denver.driver$post_date) %in% c(4, 5, 6), 1, 0)
denver.driver$quarter3 <- ifelse(month(denver.driver$post_date) %in% c(7, 8, 9), 1, 0) 
denver.driver$quarter4 <- ifelse(month(denver.driver$post_date) %in% c(10, 11, 12), 1, 0)
```

```{r}
denver.driver$quarter <- denver.driver$quarter1 * 1 + denver.driver$quarter2*2 + denver.driver$quarter3*3 + denver.driver$quarter4*4
denver.driver$quarter <- as.factor(denver.driver$quarter)
denver.driver$salary.bucket <- ifelse(denver.driver$salary < 30000, "Low", ifelse(denver.driver$salary < 50000, "Medium", "High"))
```

```{r}
library(MASS)
denver.driver$salary <-(denver.driver$salary-mean(denver.driver$salary))/sqrt(var(denver.driver$salary))
sum(denver.driver$commercial * denver.driver$training * denver.driver$quarter1 * as.factor(denver.driver$salary.bucket))
norm.salary = (denver.driver$salary - mean(denver.driver$salary))/sd(denver.driver$salary)
full.inter.model <- lm(time_to_fill ~ (norm.salary + region + safety + diploma + transportation + cust.serv + training + commercial + warehouse + quarter)^2, data = denver.driver)
step.inter.model <- stepAIC(full.inter.model, direction="both", trace=FALSE)
summary(step.inter.model)
library(car)
summary(p1 <- powerTransform(full.inter.model))
summary(m1 <- lm(bcPower(time_to_fill, p1$roundlam) ~ (norm.salary + region + safety + diploma + transportation + cust.serv + training + commercial + warehouse + quarter)^2, denver.driver))
m_best = stepAIC(m1, direction="both", trace=FALSE)
par(mfrow = c(2,2))
plot(m_best)
summary(m_best)
denver.driver2 <- denver.driver[-80,]
norm.salary = (denver.driver2$salary - mean(denver.driver2$salary))/sd(denver.driver2$salary)
summary(m1 <- lm(bcPower(time_to_fill, p1$roundlam) ~ (norm.salary + region + safety + diploma + transportation + cust.serv + training + commercial + warehouse + quarter)^2, denver.driver2))
m_best <- stepAIC(m1, direction="both", trace=FALSE)
summary(m_best)
```

```{r}
plot(step.inter.model)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=log(salary), y=log(time_to_fill))) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = log(salary), y=log(time_to_fill), color=region), se=FALSE) +
  labs(title = "Denver Driving Jobs (August 2017-July 2017)")
```

```{r}
denver.driver$quarter <- ifelse(denver.driver$quarter1 == 1, 1, ifelse(denver.driver$quarter2 == 1, 2, ifelse(denver.driver$quarter3 == 1, 3, 4)))
denver.driver$quarter <- factor(denver.driver$quarter, levels=c(3,4,1,2))
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(quarter), y = log(time_to_fill))) + 
         geom_boxplot() +
  labs(title = "Denver Driving Jobs Took Longer to Fill in 3rd, 4th Quarter 2017", x="3rd / 4th Quarter 2017                                                1st / 2nd Quarter 2018" )
```


```{r}
g1 <- ggplot(data = denver.driver, mapping = aes(x = region, y = log(time_to_fill))) + 
         geom_violin(trim=FALSE) +
  labs(title = "No Significant Difference Between Denver / Aurora / Suburbs", x="Greater Denver Area Driver Jobs")
g1 + geom_dotplot(binaxis="y", stackdir='center', dotsize=1)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(transportation), y = log(time_to_fill))) + 
         geom_dotplot(binaxis='y', stackdir='center', stackratio=1.5, dotsize=0.5) +
  labs(title = "Significant Difference Between Posts That Include Tag=(Transportation)", x="Transportation Jobs Fill Slower") + scale_x_discrete(labels=c("0" = "No Transportation", "1" = "Transportation"))
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(warehouse), y = log(time_to_fill))) + 
         geom_dotplot(binaxis='y', stackdir='center', stackratio=1.5, dotsize=0.5) +
  labs(title = "Significant Difference Between Posts That Include Tag=(Warehouse)", x="Warehouse Jobs Fill Quicker") + scale_x_discrete(labels=c("0" = "No Warehouse", "1" = "Warehouse"))
```

```{r}
denver.driver$salary.bucket <- factor(denver.driver$salary.bucket, levels=c("Low", "Medium", "High"))
ggplot(data = denver.driver, mapping = aes(x = as.factor(salary.bucket), y = log(time_to_fill))) + 
         geom_boxplot() +
  labs(title = "Denver Driving Jobs, Time-To-Fill By Salary Category", x="Under $30k Low,   $30k-$50k Medium,   Over $50k High" )
```