---
title: "PalantirDec8"
author: "Palantir"
date: "12/8/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(ggplot2)
library(tidyverse)
library(readr)
library(ggpubr)
library(reshape2)
library(lubridate)
library(MASS)

# Read data, delete repeated columns and format date
# 200,000 jobs
data_full <- read.csv('Manu_data_200000.csv')
drops <- c("job_id.1","job_id.2")
df <- data_full[, !(names(data_full) %in% drops)]
df$post_date <- as.character(df$post_date)

df$post_date <- as.Date(df[["post_date"]], "%m/%d/%Y")
df$fill_date <- as.Date(df[["fill_date"]], "%m/%d/%Y")
```



# Add Combined Statistical Areas to 11 top markets
# 2017 Estimates from Wikipedia
# NY, LA, CHI, DC, SF, DAL, BOS, ATL, DET, PHI, SJ

# df$csa <- ifelse(df$region_state=="New York-Northern New Jersey-Long Island, NY-NJ-PA MSA", 23876155, 
#                  ifelse(df$region_state=="Los Angeles-Long Beach-Santa Ana, CA MSA", 18788800,
#                  ifelse(df$region_state=="Chicago-Naperville-Joliet, IL-IN-WI MSA", 9901711, 
#                 ifelse(df$region_state=="Washington-Arlington-Alexandria, DC-VA-MD-WV MSA", 9764315, 
#                 ifelse(df$region_state=="San Francisco-Oakland-Fremont, CA MSA", 4727357, 
#                 ifelse(df$region_state=="Dallas-Fort Worth-Arlington, TX MSA", 7846293,
#                 ifelse(df$region_state=="Boston-Cambridge-Quincy, MA-NH MSA", 8233270, 
#                 ifelse(df$region_state=="Atlanta-Sandy Springs-Marietta, GA MSA", 6555956, 
#                 ifelse(df$region_state=="Detroit-Warren-Livonia, MI MSA", 5336286,
#                 ifelse(df$region_state=="Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA", 7206807,
#                 ifelse(df$region_state=="San Jose-Sunnyvale-Santa Clara, CA MSA", 1998463,NA)))))))))))

```{r}
# Aggregate Engineer and Driver jobs
# All Engineers combined (Electrical, Mechanical)
# All Drivers combined (Truck, CDL, etc)
index1 <- grep('Engineer', df$role, perl=TRUE)
df$role[index1] <- 'Engineer'
index2 <- grep('Driver', df$role, perl=TRUE)
df$role[index2] <- 'Driver'

# 181,608 jobs
# Delete jobs with time_to_fill under 2 or over 123 days
# Why? Fill time too short, might just be internal posting request
# Why? Forgotten about postings
df <- df[df$time_to_fill>2 & df$time_to_fill<123, ]

# Format month year
df$Month_Yr <- format(as.Date(df$post_date), "%Y-%m")
```

# Calculate job posting of selected roles and plot
```{r}

# group and count job id by month for Engineer and Driver
cbm <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = df[df$role=='Engineer' | df$role=='Driver',], 
          FUN = function(x){NROW(x)})

# plot over time for all regions
ggplot(data = cbm, aes(x = factor(Month_Yr), y = count, color = role)) +   
  geom_line(aes(group = role)) + geom_point() +
  xlab('Year-Month') +
  ylab('Count') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Monthly Job Postings (Nationwide, Manufacturing Vertical)")

```

# Calculate job posting of selected roles and plot
# August 2017 - July 2018
```{r}

# group and count job id by month for Engineer and Driver
cbm <- cbm[cbm$Month_Yr > "2017-07", ]

# plot over time for all regions
ggplot(data = cbm, aes(x = factor(Month_Yr), y = count, color = role)) +   
  geom_line(aes(group = role)) + geom_point() +
  xlab('Year-Month') +
  ylab('Count') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Monthly Job Postings, 8/2017 to 7/2018 (Nationwide, Manufacturing Vertical)")

```

##### 
# From this point forward, one dataset
# All dates 8/2017 to 7/2018 only
# Removed jobs with time_to_fill 0,1,2 or more than 123 days


```{r, include=FALSE}
data_full <- read_csv('Manu_data_200000.csv')
data_full<-data_full[,-c(1,17,19)]
# 200,000 observations
```

```{r}

data_full$post_date <- as.character(data_full$post_date)
data_full$fill_date <- as.character(data_full$fill_date)
data_full$post_date <- as.Date(data_full[["post_date"]], "%m/%d/%Y")
data_full$fill_date <- as.Date(data_full[["fill_date"]], "%m/%d/%Y")
data_full$Month_Yr <- format(as.Date(data_full$post_date), "%Y-%m")

# Remove time_to_fill outliers
data_full <- data_full[data_full$time_to_fill>2 & data_full$time_to_fill<123, ]
# 181,608 observations

# Remove jobs prior to August 2017
data_full <- data_full[data_full$Month_Yr > "2017-07", ]
# 177,231 observations

# str(data_full)
```

```{r}
#summary(data_full) 
```

```{r}
# time to fill distribution
summary(data_full$time_to_fill)

```

```{r}
# salary distribution
summary(data_full$salary)
```

```{r}
# Remove rows with time_to_fill = NA (4477 / 200000 ~ 2.52%)
data_full <- data_full %>% drop_na(time_to_fill)
# 172,754 observations
```

```{r}
# time to fill distribution
summary(data_full$time_to_fill)
```

# # #treat NA, tof and salary are minor numbers, removed. Mainly impute for State
# data_full <- data_full[!is.na(data_full$salary) & !is.na(data_full$time_to_fill),] 
# library(mice)
# library(VIM)
# mice_plot <- aggr(data_full, col=c('navyblue','yellow'),
#                   numbers=TRUE, sortVars=TRUE,
#                   labels=names(data_full), cex.axis=.7,
#                   gap=3, ylab=c("Missing data","Pattern"))
# temp <- data_full[,c('post_date','fill_date','time_to_fill','salary','role','state')]
# temp$state <- as.factor(temp$state)
# tempData <- mice(temp,m=5,maxit=50,meth='pmm',seed=123)
# summary(tempData)

# tempData$imp$state


# 
```{r}
 # plot trend
 tmp <- data_full %>% mutate('mydate' = format(as.Date(post_date),'%Y-%m'))%>%
   group_by(mydate) %>% summarise(num = n())
 for(ii in 1:nrow(tmp)){
   tmp$mydate[ii] <- paste(tmp$mydate[ii],'-01',sep = '')
 }
 tmp$mydate <- as.Date(tmp$mydate,'%Y-%m-%d')
 ggplot(data = tmp, aes(x = mydate, y = num)) + geom_line() 
 # It appears that after 2017-07 the number of jobs exploded
```

```{r}
# check if engineers pay and time to fill differ in groups
data_full$role<- as.character(data_full$role)
unique_engineer <- unique(data_full$role)[grep('Engineer',unique(data_full$role))]
all_engineers <- data_full[data_full$role %in% unique_engineer,]
sum_engineer <- all_engineers %>% 
  group_by(role) %>% 
  summarise(num = n(), 
  min_sal = min(salary,na.rm = T),
  mean_sal = mean(salary,na.rm = T), 
  median_sal = median(salary,na.rm = T),
  max_sal = max(salary,na.rm = T), 
  std_val = sd(salary,na.rm = T),
  min_fill = min(time_to_fill,na.rm = T),
  mean_fill = mean(time_to_fill,na.rm = T), 
  median_fill = median(time_to_fill,na.rm = T),
  max_fill = max(time_to_fill,na.rm = T), 
  std_fill = sd(time_to_fill,na.rm = T)) %>% 
  arrange(desc(num))
```



```{r}
### For presentation (Joe)

all_engineers1 <- data_full[data_full$role %in% unique_engineer,]
sum_engineer1 <- all_engineers1 %>% group_by(role) %>% summarise(
  num = n(), 
  median_sal = median(salary,na.rm = T),
  median_fill = median(time_to_fill,na.rm = T),
  std_fill = sd(time_to_fill,na.rm = T)) %>% 
  arrange(desc(num))

```


```{r}
## Boxplot for salary, all types of engineers

ggboxplot(all_engineers, x = "role", y = "salary", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# trend of the ten groups salary
ggline(all_engineers, x = "role", y = "salary", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```


```{r}
# Filter out the extreme groups and values
eng.large.role <- sum_engineer$role[sum_engineer$num>(nrow(all_engineers)*0.01)] 
eng.large.role
```



```{r}
all_engineers.filterd <- all_engineers[all_engineers$role %in% eng.large.role, ]
summary(all_engineers.filterd$salary)
length(all_engineers.filterd$salary)
```

```{r}
summary(all_engineers.filterd$time_to_fill)
length(all_engineers.filterd$time_to_fill)

```

```{r}
## get the 80% tiles of salary
eng.tiles <- ntile(all_engineers.filterd$salary,10)
```


```{r}
all_engineers.filterd <- all_engineers.filterd[eng.tiles>=2 & eng.tiles <=8,]
table(all_engineers.filterd$role) %>% sort(decreasing = T)
```

```{r}
require(scales)
```

```{r}
all_engineers.filterd1 <- all_engineers.filterd[!is.na(all_engineers.filterd$role), 
                                                ]
ggboxplot(all_engineers.filterd1, x = "role", y = "salary", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))+
  scale_y_continuous(labels = comma)
```


```{r}
# trend of the ten groups salary
ggline(all_engineers.filterd, x = "role", y = "salary", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# boxplots of ten groups time to fill
ggboxplot(all_engineers, x = "role", y = "time_to_fill", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```


```{r}
# trend of the ten groups salary
ggline(all_engineers, x = "role", y = "time_to_fill", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```
```{r}
# Filter out the extreme groups and values
eng.large.role <- sum_engineer$role[sum_engineer$num>(nrow(all_engineers)*0.01)] 
all_engineers.filterd <- all_engineers[all_engineers$role %in% eng.large.role, ]
summary(all_engineers.filterd$time_to_fill)
```
```{r}
## get the 80% tiles of salary
eng.tiles <- ntile(all_engineers.filterd$time_to_fill,10)
all_engineers.filterd <- all_engineers.filterd[eng.tiles>=2 & eng.tiles <=9,]
table(all_engineers.filterd$role) %>% sort(decreasing = T)
```
```{r}
## plot again
ggboxplot(all_engineers.filterd, x = "role", y = "time_to_fill", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```
```{r}
# trend of the ten groups
ggline(all_engineers.filterd, x = "role", y = "time_to_fill", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) # distribution looked similar
```
```{r}
## check if all engineering/drivers role following similar skill distribution and total count
# get data frame contains each role and its top 5 tags
eng.tags <- data_full %>% 
  filter(role %in% unique_engineer) %>% 
  separate_rows(.,tags,sep = ',') %>% 
  filter(!is.na(tags) & tags != '') %>% 
  group_by(role,tags) %>% 
  summarise(num = n()) %>% 
  top_n(10,num) %>% 
  arrange(role,desc(num)) %>% ungroup()

sort(table(eng.tags$tags),decreasing = T)
```

```{r}
eng.tags.max.num <- eng.tags %>% 
  group_by(role) %>% 
  summarise(max_num = max(num)) 
eng.tags.max.num <- eng.tags.max.num[eng.tags.max.num$max_num > 100,]
eng.tags <- inner_join(eng.tags,eng.tags.max.num, by = 'role')
eng.tags$perc <- eng.tags$num/eng.tags$max_num
```

```{r}
# reshape it to calculate similarity

wide.eng.tags <- dcast(eng.tags,role~tags,value.var = 'perc' )
wide.eng.tags[is.na(wide.eng.tags)] <- 0
```

```{r}
# quick clustering
d <- dist(wide.eng.tags,method = 'euclidean') # distance matrix
d # distance similar, safe to combine engineers together

```


```{r}
## DRIVERS
unique_drivers <- unique(data_full$role)[grep('Driver',unique(data_full$role))]
all_drivers <- data_full[data_full$role %in% unique_drivers,]
sum_drivers <- all_drivers %>% group_by(role) %>% summarise(
  num = n(),  
  min_sal = min(salary,na.rm = T),
  mean_sal = mean(salary,na.rm = T), 
  median_sal = median(salary,na.rm = T),
  max_sal = max(salary,na.rm = T), 
  std_val = sd(salary,na.rm = T),
  min_fill = min(time_to_fill,na.rm = T),
  mean_fill = mean(time_to_fill,na.rm = T), 
  median_fill = median(time_to_fill,na.rm = T),
  max_fill = max(time_to_fill,na.rm = T), 
  std_fill = sd(time_to_fill,na.rm = T)) %>% 
  arrange(desc(num))
```

```{r}
### For presentation (Joe)

all_drivers1 <- data_full[data_full$role %in% unique_drivers,]
sum_drivers1 <- all_drivers1 %>% group_by(role) %>% summarise(
  num = n(),  
  median_sal = median(salary,na.rm = T),
  median_fill = median(time_to_fill,na.rm = T),
  std_fill = sd(time_to_fill,na.rm = T)) %>% 
  arrange(desc(num))

```


```{r}
## Boxplot for this
# boxplots of ten groups salary

ggboxplot(all_drivers, x = "role", y = "salary", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) +
  scale_y_continuous(labels = comma)
```

```{r}
# trend of the ten groups salary
ggline(all_drivers, x = "role", y = "salary", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# Filter out the extreme groups and values
dri.large.role <- sum_drivers$role[sum_drivers$num>(nrow(all_drivers)*0.01)] 
all_drivers.filterd <- all_drivers[all_drivers$role %in% dri.large.role, ]
summary(all_drivers.filterd$salary)
length(all_drivers.filterd$salary)
```


```{r}
## get the 80% tiles of salary
dri.tiles <- ntile(all_drivers.filterd$salary,10)
all_drivers.filterd <- all_drivers.filterd[dri.tiles>=2 & dri.tiles <=9,]
table(all_drivers.filterd$role) %>% sort(decreasing = T)
```

```{r}
all_drivers.filterd1 <- all_drivers.filterd[!is.na(all_drivers.filterd$role), 
                                                ]
ggboxplot(all_drivers.filterd1, x = "role", y = "salary", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) +
  scale_y_continuous(labels = comma)
```

```{r}
# trend of the ten groups salary
ggline(all_drivers.filterd, x = "role", y = "salary", color = "role", 
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```
```{r}
# boxplots of ten groups time to fill
ggboxplot(all_drivers, x = "role", y = "time_to_fill", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# trend of the ten groups salary
ggline(all_drivers, x = "role", y = "time_to_fill", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# Filter out the extreme groups and values
dri.large.role <- sum_drivers$role[sum_drivers$num>(nrow(all_drivers)*0.01)] 
all_drivers.filterd <- all_drivers[all_drivers$role %in% dri.large.role, ]
summary(all_drivers.filterd$time_to_fill)
```

```{r}
## get the 80% tiles of salary
dri.tiles <- ntile(all_drivers.filterd$time_to_fill,10)
#all_drivers.filterd <- all_drivers.filterd[dri.tiles>=2 & dri.tiles <=8,]
table(all_drivers.filterd$role) %>% sort(decreasing = T)
```

```{r}
## plot again
ggboxplot(all_drivers.filterd, x = "role", y = "time_to_fill", fill = 'role',
          palette = 'magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7))
```

```{r}
# trend of the ten groups
ggline(all_drivers.filterd, x = "role", y = "time_to_fill", color = "role",
       add = c("mean_se",'violin'),
       palette ='magma')+theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) 
# distribution looked similar
```

```{r}
## check if all enginnering/drivers role following similar skill distribution and total count
# get data frame contains each role and its top 5 tags
dri.tags <- data_full %>% filter(role %in% unique_drivers) %>% 
  separate_rows(.,tags,sep = ',') %>% filter(!is.na(tags) & tags != '') %>% 
  group_by(role,tags) %>% summarise(num = n()) %>% top_n(10,num) %>% 
  arrange(role,desc(num)) %>% ungroup()
sort(table(dri.tags$tags),decreasing = T)
```

```{r}
dri.tags.max.num <- dri.tags %>% group_by(role) %>% summarise(max_num = max(num)) 
dri.tags.max.num <- dri.tags.max.num[dri.tags.max.num$max_num > 100,]
dri.tags <- inner_join(dri.tags,dri.tags.max.num, by = 'role')
dri.tags$perc <- dri.tags$num/dri.tags$max_num
```

```{r}
# reshape it to calculate similarity
wide.dri.tags <- dcast(dri.tags,role~tags,value.var = 'perc' )
wide.dri.tags[is.na(wide.dri.tags)] <- 0
# quick clustering
d <- dist(wide.dri.tags,method = 'euclidean') # distance matrix
d # distance similar, safe to combine drivers together
```

```{r}
#combine roles 
unique_engineer <- unique_engineer[unique_engineer!= 'Engineer']
unique_drivers <- unique_drivers[unique_drivers!= 'Driver']
data_full$role[data_full$role %in% unique_engineer] <- 'Engineer'
data_full$role[data_full$role %in% unique_drivers] <- 'Driver'
```


```{r}
#Engineering/driver job demand trend (number of engineer/driver job over total jobs that month)
tmp <- data_full %>% 
  mutate('mydate' = format(as.Date(post_date),'%Y-%m')) %>% 
  group_by(mydate,role) %>% 
  summarise(num = n())

x <- (tmp %>% filter(role %in% c('Engineer','Driver')) %>% 
        group_by(mydate,role)%>%
        summarise(num = sum(num)))

y <- (tmp %>% group_by(mydate) %>% summarise(num = sum(num)))

tmp <- merge(x,y, by = 'mydate')

tmp$rate <- tmp$num.x/tmp$num.y

for(ii in 1:nrow(tmp)){
  tmp$mydate[ii] <- paste(tmp$mydate[ii],'-01',sep = '')
}

tmp$mydate <- as.Date(tmp$mydate,'%Y-%m-%d')
ggplot(data = tmp, aes(x = mydate, y = rate, color = role)) + geom_line() 
# after 2017 07 much stabler.
```



```{r}
#summerise rates
summary(tmp[tmp$role == 'Engineer','rate'])
summary(tmp[tmp$role == 'Driver','rate'])
```


```{r}
## Split to Drivers and Engineers
# check time to fill and salary distribution
data_engineer <- data_full[data_full$role =='Engineer',]
data_driver <- data_full[data_full$role == 'Driver',]
```

```{r}
summary(data_engineer$time_to_fill)
summary(data_engineer$salary)
summary(data_driver$time_to_fill)
summary(data_driver$salary)
```

```{r}
ggboxplot(rbind(data_engineer,data_driver), x = "role", y = "time_to_fill", fill = 'role', palette = 'magma') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) 
# although mean close, engineers seems to take longer to fill
```

```{r}
ggboxplot(rbind(data_engineer,data_driver), x = "role", y = "salary", fill = 'role',
          palette = 'magma') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1,size = 7)) 
# engineers salary higher on average
```

```{r}
# kept only 80% inner quantile of all salaries
eng.tiles <- ntile(data_engineer$salary,10)
data_engineer <- data_engineer[eng.tiles>=2 & eng.tiles <=9,] #much more normal
```

```{r}
dri.tiles <- ntile(data_driver$salary,10)
data_driver <- data_driver[dri.tiles>=2 & dri.tiles <=9,] #much more normal
```


```{r}
## look at time_to_fill, (not any longer, already removed earlier under 3 days and over 123 days) ***keep inner 80%. 
sort(table(data_engineer$time_to_fill),decreasing = T) 
#eng.time.tiles <- ntile(data_engineer$time_to_fill,10)
##data_engineer$time_to_fill[eng.time.tiles>=2 & eng.time.tiles <=9] %>% summary() 
data_engineer$time_to_fill%>% summary()
# 4 -115


```

```{r}
#data_engineer <- data_engineer[eng.time.tiles>=2 & eng.time.tiles <=9,] # 26148 rows
sort(table(data_driver$time_to_fill),decreasing = T) 
#dri.time.tiles <- ntile(data_driver$time_to_fill,10)
#data_driver$time_to_fill[dri.time.tiles>=2 & dri.time.tiles <=9] %>% summary() # 3 -94
#data_driver <- data_driver[dri.time.tiles>=2 & dri.time.tiles <=9,] # 21130 row
# distribution of median salary and time fill over time
data_driver$time_to_fill %>% summary()
```

# ```{r}
# tmp <- data_full %>% mutate('mydate' = format(as.Date(post_date),'%Y-%m')) %>% 
#   select(mydate,role,salary) %>% filter(role %in% c('Engineer','Driver')) %>% 
#   group_by(mydate,role) %>% summarise(med_salary = median(salary,na.rm = T))
# 
# 
# for(ii in 1:nrow(tmp)){
#   tmp$mydate[ii] <- paste(tmp$mydate[ii],'-01',sep = '')
# }
# tmp$mydate <- as.Date(tmp$mydate,'%Y-%m-%d')
# ggplot(data = tmp, aes(x = mydate, y = med_salary, color = role)) + geom_line()
# ```


# ```{r}
# tmp_2 <- tmp
# tmp_2$med_salary[tmp_2$role == 'Driver'] <-tmp_2$med_salary[tmp_2$role == 'Driver'] + 
#   (median(data_engineer$salary,na.rm = T)-median(data_driver$salary,na.rm = T))
# ggplot(data = tmp_2, aes(x = mydate, y = med_salary, color = role)) + geom_line()
# ```



```{r}
#data_full$tags <- as.character(data_full$tags)
#data_engineer$tags <- as.character(data_engineer$tags)
#data_driver$tags <- as.character(data_driver$tags)
```



```{r}
## write the csvs 
str(data_engineer)
write_csv(data_full,'/nfs/home/jjk5642/z/msia-400-project1/kupe/data_full.csv')
write_csv(data_engineer,'/nfs/home/jjk5642/z/msia-400-project1/kupe/data_engineer.csv')
write_csv(data_driver,'/nfs/home/jjk5642/z/msia-400-project1/kupe/data_driver.csv')
```

# time series analysis
# calculate and plot ratio of job count
```{r}

# total job count by month_year
total_job <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full, 
          FUN = function(x){NROW(x)})

# job count by month_year of engineer and driver
cbm <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Engineer'|data_full$role=='Driver'),], 
          FUN = function(x){NROW(x)})

# get ratio of job counts (rescale)
cbm <- merge(cbm, total_job, by = 'Month_Yr', all = TRUE)
cbm$ratio <- cbm$count.x / cbm$count.y

# plot over time for all regions
# both driver and engineer
ggplot(data = cbm, aes(x = factor(Month_Yr), y = ratio, color= role)) +
  geom_line(aes(group=role)) + geom_point() +
  xlab('Year-Month') +
  ylab('Ratio of Job Posting') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Proportion of Manufacturing Vertical, 8/2017 to 7/2018 (Nationwide)")
```

# ```{r}
# # plot ratio for engineer
# ggplot(data = cbm[cbm$role=='Engineer',], aes(x = factor(Month_Yr), y = ratio, color~role)) +
# geom_line(aes(group=role)) + geom_point() + 
#   ggtitle('Engineer Jobs as a Proportion of All Manufacturing Jobs') +
#   xlab('Year-Month') +
#   ylab('Ratio of Job Posting') +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))
# ```
```{r}

# total job count by month_year
total_job_sj <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="San Jose-Sunnyvale-Santa Clara, CA MSA", ], 
          FUN = function(x){NROW(x)})

total_job_sf <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="San Francisco-Oakland-Fremont, CA MSA", ], 
          FUN = function(x){NROW(x)})

# job count by month_year of engineer
cbm_sj <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Engineer' & data_full$region_state=="San Jose-Sunnyvale-Santa Clara, CA MSA"),], 
          FUN = function(x){NROW(x)})

cbm_sf <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Engineer' & data_full$region_state=="San Francisco-Oakland-Fremont, CA MSA"),], 
          FUN = function(x){NROW(x)})

# get ratio of job counts (rescale)
cbm_sj <- merge(cbm_sj, total_job_sj, by = 'Month_Yr', all = TRUE)
cbm_sj$ratio <- cbm_sj$count.x / cbm_sj$count.y
cbm_sj$region = "San Jose"

cbm_sf <- merge(cbm_sf, total_job_sf, by = 'Month_Yr', all = TRUE)
cbm_sf$ratio <- cbm_sf$count.x / cbm_sf$count.y
cbm_sf$region = "San Francisco"

cbm <- rbind(cbm_sj, cbm_sf)
# plot over time for all regions
# both driver and engineer


ggplot(data=cbm, aes(x = factor(Month_Yr), y = ratio, color=region)) +
  geom_line(aes(group=region)) + geom_point() +
  xlab('Year-Month') +
  ylab('Ratio of Job Posting') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Engineering Proportion of Manufacturing Vertical, 8/2017 to 7/2018")
```

```{r}

# total job count by month_year
total_job_det <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="Detroit-Warren-Livonia, MI MSA", ], 
          FUN = function(x){NROW(x)})

total_job_chi <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="Chicago-Naperville-Joliet, IL-IN-WI MSA", ], 
          FUN = function(x){NROW(x)})

total_job_atl <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="Atlanta-Sandy Springs-Marietta, GA MSA", ], 
          FUN = function(x){NROW(x)})

total_job_phi <- aggregate(cbind(count = job_id) ~ Month_Yr, 
          data = data_full[data_full$region_state=="Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA", ], 
          FUN = function(x){NROW(x)})

# job count by month_year for driver 
cbm_det <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Driver' & data_full$region_state=="Detroit-Warren-Livonia, MI MSA"),], 
          FUN = function(x){NROW(x)})

cbm_chi <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Driver' & data_full$region_state=="Chicago-Naperville-Joliet, IL-IN-WI MSA"),], 
          FUN = function(x){NROW(x)})

cbm_atl <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Driver' & data_full$region_state=="Atlanta-Sandy Springs-Marietta, GA MSA"),], 
          FUN = function(x){NROW(x)})

cbm_phi <- aggregate(cbind(count = job_id) ~ Month_Yr+role, 
          data = data_full[which(data_full$role=='Driver' & data_full$region_state=="Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA"),], 
          FUN = function(x){NROW(x)})

# get ratio of job counts (rescale)
cbm_det <- merge(cbm_det, total_job_det, by = 'Month_Yr', all = TRUE)
cbm_det$ratio <- cbm_det$count.x / cbm_det$count.y
cbm_det$region = "Detroit"

cbm_chi <- merge(cbm_chi, total_job_chi, by = 'Month_Yr', all = TRUE)
cbm_chi$ratio <- cbm_chi$count.x / cbm_chi$count.y
cbm_chi$region = "Chicago"

cbm_atl <- merge(cbm_atl, total_job_atl, by = 'Month_Yr', all = TRUE)
cbm_atl$ratio <- cbm_atl$count.x / cbm_atl$count.y
cbm_atl$region = "Atlanta"

cbm_phi <- merge(cbm_phi, total_job_phi, by = 'Month_Yr', all = TRUE)
cbm_phi$ratio <- cbm_phi$count.x / cbm_phi$count.y
cbm_phi$region = "Philly"

cbm <- rbind(cbm_det, cbm_chi, cbm_atl, cbm_phi)
# plot over time for all regions
# both driver and engineer


ggplot(data=cbm, aes(x = factor(Month_Yr), y = ratio, color=region)) +
  geom_line(aes(group=region)) + geom_point() +
  xlab('Year-Month') +
  ylab('Ratio of Job Posting') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Driving Proportion of Manufacturing Vertical, 8/2017 to 7/2018")
```


# ```{r}
# # plot ratio for driver
# ggplot(data = cbm[cbm$role=='Driver',], aes(x = factor(Month_Yr), y = ratio, color~role)) +
#   geom_line(aes(group=role)) + geom_point() +
#   ggtitle('Driver') + 
#   xlab('Year-Month') +
#   ylab('Ratio of Job Posting') + 
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))
# ```

# cross market analysis
# ratio of Engineer by region_state (same code for driver: replacing Engineer with Driver)
```{r}

# total job count by month_year and region_state
total_job_state <- aggregate(cbind(count = job_id) ~ Month_Yr+region_state, 
          data = data_full, 
          FUN = function(x){NROW(x)})
total_job_state <- total_job_state[total_job_state$Month_Yr > "2017-07", ]

# job count by month_year and region state of engineer
cbm_state <- aggregate(cbind(count = job_id) ~ Month_Yr+region_state, 
          data = data_full[which(data_full$role=='Engineer'),], 
          FUN = function(x){NROW(x)})
cbm_state <- cbm_state[cbm_state$Month_Yr > "2017-07", ]

# get ratio of job counts (rescale)
cbm_state <- merge(cbm_state, total_job_state, by = c('Month_Yr', 'region_state'), all = TRUE)
cbm_state[is.na(cbm_state)] <- 0
cbm_state$ratio <- cbm_state$count.x / cbm_state$count.y

# find top 5 regions having largest job count of engineer
cbs_e <- aggregate(cbind(count = job_id) ~ region_state,
                 data = data_full[data_full$role=='Engineer',],
                 FUN = function(x){NROW(x)})
cbs_e <- cbs_e[order(-cbs_e$count), c(1,2)]
top <- head(cbs_e[, 1],5)
#print(head(cbs_e,10))

# find data on top5 region 
cbm_state_top <- cbm_state[is.element(cbm_state$region_state, top),]
cbm_state_top <- cbm_state_top[cbm_state_top$Month_Yr > "2017-07", ]

# plot ratio of time by state
ggplot(data = cbm_state_top, aes(x = factor(Month_Yr), y = ratio, color = region_state)) +   
  geom_line(aes(group = region_state)) + geom_point() +
  ggtitle('Engineer Job Postings, Top Five Metro Regions') +
  xlab('Year-Month') +
  ylab('Ratio of Job Posting') +
  theme(legend.position='right', 
        legend.text=element_text(size=7),
        axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}

# total job count by month_year and region_state
total_job_state <- aggregate(cbind(count = job_id) ~ Month_Yr+region_state, 
          data = data_full, 
          FUN = function(x){NROW(x)})
total_job_state <- total_job_state[total_job_state$Month_Yr > "2017-07", ]

# job count by month_year and region state of driver
cbm_state <- aggregate(cbind(count = job_id) ~ Month_Yr+region_state, 
          data = data_full[which(data_full$role=='Driver'),], 
          FUN = function(x){NROW(x)})
cbm_state <- cbm_state[cbm_state$Month_Yr > "2017-07", ]

# get ratio of job counts (rescale)
cbm_state <- merge(cbm_state, total_job_state, by = c('Month_Yr', 'region_state'), all = TRUE)
cbm_state[is.na(cbm_state)] <- 0
cbm_state$ratio <- cbm_state$count.x / cbm_state$count.y

# find top 5 regions having largest job count of engineer
cbs_e <- aggregate(cbind(count = job_id) ~ region_state,
                 data = data_full[data_full$role=='Driver',],
                 FUN = function(x){NROW(x)})
cbs_e <- cbs_e[order(-cbs_e$count), c(1,2)]
top <- head(cbs_e[, 1],5)
#print(head(cbs_e,10))

# find data on top5 region 
cbm_state_top <- cbm_state[is.element(cbm_state$region_state, top),]
cbm_state_top <- cbm_state_top[cbm_state_top$Month_Yr > "2017-07", ]

# plot ratio of time by state
ggplot(data = cbm_state_top, aes(x = factor(Month_Yr), y = ratio, color = region_state)) +   
  geom_line(aes(group = region_state)) + geom_point() +
  ggtitle('Driver Job Postings, Top Five Metro Regions') +
  xlab('Year-Month') +
  ylab('Ratio of Job Posting') +
  theme(legend.position='right', 
        legend.text=element_text(size=5),
        axis.text.x = element_text(angle = 90, hjust = 1))



```


# Just look at Denver Drivers as an example to simplify
# Why?  Noise in the job market - many types of jobs, many differing regions

```{r}
# 412 jobs
denver.driver <- filter(data_full, role %in% c("Truck Driver", "Driver", "CDL Driver"), region_state=="Denver-Aurora, CO MSA")
# Remove jobs only posted for 0 or 1 day
# 387 jobs, NA removed by default
denver.driver <- filter(denver.driver, time_to_fill > 1)
# Push the job id to the end
denver.driver$job_id_end <- denver.driver$job_id
denver.driver <- denver.driver[-c(1)]
denver.driver$region <- ifelse(denver.driver$city == "Denver", "Denver", ifelse(denver.driver$city == "Aurora", "Aurora", "Other"))
denver.driver$region <- as.factor(denver.driver$region)
denver.driver$region <- relevel(denver.driver$region, "Other")
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=salary, y=time_to_fill)) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = salary, y=time_to_fill, color=region), se=FALSE)
```

```{r}
# Try to predict time to fill based on salary, region = ("Other", "Denver", "Aurora")
den.driver.model <- lm(time_to_fill ~ salary + region, data=denver.driver)
summary(den.driver.model)
```

```{r}
# Remove extra high salaries (Many jobs have "accounting" in tags)
# 373 jobs, NA removed by default
denver.driver <- filter(denver.driver, salary < 100000)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=salary, y=time_to_fill)) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = salary, y=time_to_fill, color=region), se=FALSE) +
  labs(title="Denver Driving Job Market (August 2017-July 2018")
```

```{r}
# Try to predict time to fill based on salary, region = ("Other", "Denver", "Aurora")
den.driver.model <- lm(time_to_fill ~ salary + region, data=denver.driver)
summary(den.driver.model)
# Inside / Outside Denver or Aurora or suberbs does not matter
```

# There was barely a relationship between "time_to_fill" and "salary" or "region = Denver, Aurora, Other" and when outliers removed, things are even worse. 

# Add indicator tags to determine if there are any relationships with the top driving tags:
# "Safety", "High School Diploma", "Training", "Transportation", "Customer Service", "Commercial", "Warehouse"

```{r}
denver.driver$safety <- as.numeric(str_detect(denver.driver$tags, "Safety"))
denver.driver$diploma <- as.numeric(str_detect(denver.driver$tags, "High School Diploma"))
denver.driver$training <- as.numeric(str_detect(denver.driver$tags, "Training"))
denver.driver$transportation <- as.numeric(str_detect(denver.driver$tags, "Transportation"))
denver.driver$cust.serv <- as.numeric(str_detect(denver.driver$tags, "Customer Service"))
denver.driver$commercial <- as.numeric(str_detect(denver.driver$tags, "Commercial"))
denver.driver$warehouse <- as.numeric(str_detect(denver.driver$tags, "Warehouse"))
```

```{r}
denver.driver$post_date<-as.Date(denver.driver$post_date,"%m/%d/%Y")
denver.driver$fill_date<-as.Date(denver.driver$fill_date,"%m/%d/%Y")
```

```{r}
# Average time to fill driving jobs in Denver
# Month 1 = August 2017
# Month 12 = July 2018

```

```{r}
# Driver job postings by week in Denver
denver.driver %>%
  count(week = floor_date(post_date, "week")) %>%
  ggplot(aes(week, n)) + 
  geom_line() +
  labs(title = "Denver Driving Jobs (August 2017-July 2017), Job Posts Per Week")
```

```{r}
denver.driver$quarter1 <- ifelse(month(denver.driver$post_date) %in% c(1, 2, 3), 1, 0) 
denver.driver$quarter2 <- ifelse(month(denver.driver$post_date) %in% c(4, 5, 6), 1, 0)
denver.driver$quarter3 <- ifelse(month(denver.driver$post_date) %in% c(7, 8, 9), 1, 0) 
denver.driver$quarter4 <- ifelse(month(denver.driver$post_date) %in% c(10, 11, 12), 1, 0)
```

```{r}
denver.driver$salary.bucket <- ifelse(denver.driver$salary < 30000, "Low", ifelse(denver.driver$salary < 50000, "Medium", "High"))
```

```{r}

full.den.driver.model <- lm(log(time_to_fill) ~ log(salary) + region + safety + diploma + transportation + cust.serv + training + commercial + warehouse + quarter1 + quarter2 + quarter3 + salary.bucket, data=denver.driver)
summary(full.den.driver.model)
```

```{r}
step.denver.driver.model <- stepAIC(full.den.driver.model, direction="both", trace=FALSE)
summary(step.denver.driver.model)
```

```{r}
plot(step.denver.driver.model)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x=log(salary), y=log(time_to_fill))) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(mapping = aes(x = log(salary), y=log(time_to_fill), color=region), se=FALSE) +
  labs(title = "Denver Driving Jobs (August 2017-July 2017)")
```

```{r}
denver.driver$quarter <- ifelse(denver.driver$quarter1 == 1, 1, ifelse(denver.driver$quarter2 == 1, 2, ifelse(denver.driver$quarter3 == 1, 3, 4)))
denver.driver$quarter <- factor(denver.driver$quarter, levels=c(3,4,1,2))
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(quarter), y = log(time_to_fill))) + 
         geom_boxplot() +
  labs(title = "Denver Driving Jobs Took Longer to Fill in 3rd, 4th Quarter 2017", x="3rd / 4th Quarter 2017                                                1st / 2nd Quarter 2018" )
```


```{r}
g1 <- ggplot(data = denver.driver, mapping = aes(x = region, y = log(time_to_fill))) + 
         geom_violin(trim=FALSE) +
  labs(title = "No Significant Difference Between Denver / Aurora / Suburbs", x="Greater Denver Area Driver Jobs")
g1 + geom_dotplot(binaxis="y", stackdir='center', dotsize=1)
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(transportation), y = log(time_to_fill))) + 
         geom_dotplot(binaxis='y', stackdir='center', stackratio=1.5, dotsize=0.5) +
  labs(title = "Significant Difference Between Posts That Include Tag=(Transportation)", x="Transportation Jobs Fill Slower") + scale_x_discrete(labels=c("0" = "No Transportation", "1" = "Transportation"))
```

```{r}
ggplot(data = denver.driver, mapping = aes(x = as.factor(warehouse), y = log(time_to_fill))) + 
         geom_dotplot(binaxis='y', stackdir='center', stackratio=1.5, dotsize=0.5) +
  labs(title = "Significant Difference Between Posts That Include Tag=(Warehouse)", x="Warehouse Jobs Fill Quicker") + scale_x_discrete(labels=c("0" = "No Warehouse", "1" = "Warehouse"))
```

```{r}
denver.driver$salary.bucket <- factor(denver.driver$salary.bucket, levels=c("Low", "Medium", "High"))
ggplot(data = denver.driver, mapping = aes(x = as.factor(salary.bucket), y = log(time_to_fill))) + 
         geom_boxplot() +
  labs(title = "Denver Driving Jobs, Time-To-Fill By Salary Category", x="Under $30k Low,   $30k-$50k Medium,   Over $50k High" )
```
